### 2018-06-02 draft.Rmd
Samira edited text

Almost all speakers of Andi and other Andic languages speak Avar and Russian in addition to their mother tongue. Avar was the major local lingua franca in this area before the introduction of Russian and, to a certain extent, still is. Speakers of Andic (as well as speakers of Tsezic) languages use it to communicate with each other and with local Avars. The language is also taught in school as a subject called "native language" (*rodnoj jazyk* in Russian). For speakers of Andi, Avar is the language of literacy. Before the introduction of modern technology like computers and cellphones, they would write letters to each other in Avar. Currently, they use the extended Cyrillic script of Avar (featuring, among other things, special diacritics for ejective consonants) to write their own language. Russian was introduced as the language of administration and education in the first half of the 20th century and subsequently has strengthened its position throughout the republic, cf. *ССЫЛКА НА ДОБРУШИНУ И КОЖУХАР*. Andi speakers use it to communicate with people from other regions of Daghestan where Avar is not spoken. It is also used for communication by speakers of Andi dialects that belong to different dialect groups.
	As a result of this multilingualism, the Andi lexicon contains multiple layers of borrowings from different languages and different periods. Besides loans from Avar and Russian this includes religious terminology from Arabic, and cultural borrowings from Arabic, Persian and Kumyk (a local Turkic language), some of which were evidently borrowed through Avar. The names of all the technical innovations of the 20th and 21st centuries were taken from Russian with little to no phonetic adaptation. In Russian these terms were mostly borrowed and adapted from English. The lack of phonetic adaptation in more recent Russian loans is connected to the increasingly important role of the Russian language in Daghestan. Earlier loans were subjected to some phonetic change, cf. *kempeti* (*kanfeti*, 'sweets'), *tʃajnik'i* (*tʃajnik* 'kettle'), and *ʃːiɡolati* (*ʃːakalat* 'chocolate' [@salimov10: 43], as opposed to more recent words (see Section 4). Borrowed words often have some phonological, orthographic, morphological or syntactic features in the donor language which are incompatible with the system of the recipient language. In this paper, we explore how recent Russian borrowings are integrated in the Andi system of noun classes (see related studies [gagliardi14], [stolz09], on *ЗДЕСЬ НАДО УТОЧНИТЬ, О ЧЕМ ЭТИ СТАТЬИ, А ТО ПОХОЖЕ ЧТО ОНИ НАПИСАЛИ ИМЕННО ПРО КЛАССЫ В АНДИЙСКОМ. КАК ОНИ RELATED К НАМ?*).

@@

[the guest m/f arrived (example)]

@@

  In the descriptive grammar by [@suleymanov57] (Rikvani dialect), a dictionary of 1236 nouns is provided, grouped by semantic category and noun class. The vast majority of these nouns belongs to the inanimate 1 class (678). There are 53 masculine nouns, 38 feminine, and 46 ambiguous human nouns. A total of 29 nouns is classified as ambiguous between masculine, feminine and animate (e.g. *motʃi* 'child'). The animate class consists of 109 nouns. There are 276 nouns which belong to inanimate 2, and the third inanimate class contains just seven insects. As the first inanimate class comprises half of the noun vocabulary, it covers a large number of rather general semantic categories, including body parts, plants, household objects and abstract concepts [@suleymanov57:138-151]. Unfortunately, the second inanimate class seems to cover largely the same categories, just different parts of it. For example, *rel'a* 'hand' belongs to the first inanimate class, while *гъажу* 'arm' belongs to the second. Few of the categories are exclusive as a whole to either inanimate 1 or inanimate 2, with the exception of a few highly specialized categories belonging to inanimate 2: religious objects and moral concepts (e.g. prayer, tax, amulet); paper objects and documents (book, letter, newspaper, various legal documents); concepts of time (hour, day, night). Curiously, female clothing and jewelry is mostly inanimate 2, while male clothing predominantly belongs to inanimate 1.
  The situation is a bit different in Gagatli, as described in [@salimov10: 48-]. In Gagatli, the feminine class contains a handful of abstract concepts that belong to inanimate 2 in Rikvani (e.g. evening). Otherwise, the description of the masculine and animate classes, as well as that of the inanimate 1 and 2 classes is highly similar. The insects distinguished by the third inanimate class in Rikvani are grouped with inanimate 2 in Gagatli. For both dialects liquids and water reservoirs tend to be assigned the inanimate 2 class. For Zilo this largely hold, although *inχi* 'large river' belongs to inanimate 1. 
    A few remarks on diachrony are in order here. As Table 2 illustrates, the prefixal marker of inanimate plurals is similar to the inanimate 2 marker in Andi. Three-class systems are considered reduced systems originating in a four or five class system that is more similar to what we find in contemporary Andi *ССЫЛКА НА КАКУЮ-НИБУДЬ РАБОТУ АЛЕКСЕЕВА*. As argued in [@daniel18] there is some evidence that across the family, the second inanimate class (in cases where there is such a distinction) and similar-looking plural markers, ultimately originate in a distinct class containing mass nouns, such as liquids and abstract nouns that are "number neutral". If that is the case inanimate 2 has extended its terrain considerably in contemporary Andi, as it contains many countable objects. In addition, mass nouns are not restricted to inanimate 2, for example *zar* 'ice' and *k'ːöj* 'smoke' belong to inanimate 1 at least in Rikvani. Thus, evidence from Andi does not necessarily refute the mass noun hypothesis, but it cannot support it either, as we have no knowledge of how Andi noun classes have evolved. Moreover because we have not yet succeeded in establishing semantic regularities for the synchronic level.

>>>>

Almost all speakers of Andi and other Andic languages speak Avar and Russian in addition to their mother tongue. Avar was the major local lingua franca in this area before the introduction of Russian and, to a certain extent, still is. Speakers of Andic (as well as speakers of Tsezic) languages use it to communicate with each other and with local Avars. The language is also taught in school as a subject called "native language" (*rodnoj jazyk* in Russian). For speakers of Andi, Avar is the language of literacy. Before the introduction of modern technology like computers and cellphones, they would write letters to each other in Avar. Currently, they use the extended Cyrillic script of Avar (featuring, among other things, special diacritics for ejective consonants) to write their own language. Russian was introduced as the language of administration and education in the first half of the 20th century and subsequently has strengthened its position throughout the republic, cf. [@dobrushina11]. Andi speakers use it to communicate with people from other regions of Daghestan where Avar is not spoken. It is also used for communication by speakers of Andi dialects that belong to different dialect groups.
	As a result of this multilingualism, the Andi lexicon contains multiple layers of borrowings from different languages and different periods. Besides loans from Avar and Russian this includes religious terminology from Arabic, and cultural borrowings from Arabic, Persian and Kumyk (a local Turkic language), some of which were evidently borrowed through Avar. The names of all the technical innovations of the 20th and 21st centuries were taken from Russian with little to no phonetic adaptation. In Russian these terms were mostly borrowed and adapted from English. The lack of phonetic adaptation in more recent Russian loans is connected to the increasingly important role of the Russian language in Daghestan. Earlier loans were subjected to some phonetic change, cf. *kempeti* (*kanfeti*, 'sweets'), *tʃajnik'i* (*tʃajnik* 'kettle'), and *ʃːiɡolati* (*ʃːakalat* 'chocolate' [@salimov10: 43], as opposed to more recent words (see Section 4). Borrowed words often have some phonological, orthographic, morphological or syntactic features in the donor language which are incompatible with the system of the recipient language. In this paper, we explore how recent Russian borrowings are integrated in the Andi system of noun classes (see related studies [gagliardi14], [stolz09], on *ЗДЕСЬ НАДО УТОЧНИТЬ, О ЧЕМ ЭТИ СТАТЬИ, А ТО ПОХОЖЕ ЧТО ОНИ НАПИСАЛИ ИМЕННО ПРО КЛАССЫ В АНДИЙСКОМ. КАК ОНИ RELATED К НАМ?*).

@@

* **ʃːu-w ketʃ'eqan**
good-m singer
'a good singer (male)'

* **ʃːu-j ketʃ'eqan**
good-f singer
'a good singer (female)'

@@

#### 3.1 Inanimate classes

In the descriptive grammar by [@suleymanov57] (Rikvani dialect), a dictionary of 1236 nouns is provided, grouped by semantic category and noun class. The vast majority of these nouns belongs to the inanimate 1 class (678). There are 53 masculine nouns, 38 feminine, and 46 ambiguous human nouns. A total of 29 nouns is classified as ambiguous between masculine, feminine and animate (e.g. *motʃi* 'child'). The animate class consists of 109 nouns. There are 276 nouns which belong to inanimate 2, and the third inanimate class contains just seven insects. As the first inanimate class comprises half of the noun vocabulary, it covers a large number of rather general semantic categories, including body parts, plants, household objects and abstract concepts [@suleymanov57:138-151]. Unfortunately, the second inanimate class seems to cover largely the same categories, just different parts of it. For example, *rel'a* 'hand' belongs to the first inanimate class, while *гъажу* 'arm' belongs to the second. Few of the categories are exclusive as a whole to either inanimate 1 or inanimate 2, with the exception of a few highly specialized categories belonging to inanimate 2: religious objects and moral concepts (e.g. prayer, tax, amulet); paper objects and documents (book, letter, newspaper, various legal documents); concepts of time (hour, day, night). Curiously, female clothing and jewelry is mostly inanimate 2, while male clothing predominantly belongs to inanimate 1.The situation is a bit different in Gagatli, as described in [@salimov10: 48-]. In Gagatli, the feminine class contains a handful of abstract concepts that belong to inanimate 2 in Rikvani (e.g. evening). Otherwise, the description of the masculine and animate classes, as well as that of the inanimate 1 and 2 classes is highly similar. The insects distinguished by the third inanimate class in Rikvani are grouped with inanimate 2 in Gagatli. For both dialects liquids and water reservoirs tend to be assigned the inanimate 2 class. For Zilo this largely hold, although *inχi* 'large river' belongs to inanimate 1.

#### 3.2 Origins of inanimate 2

A few remarks on diachrony are in order here. As Table 2 illustrates, the prefixal marker of inanimate plurals is similar to the inanimate 2 marker in Andi. Three-class systems are considered reduced systems originating in a four or five class system that is more similar to what we find in contemporary Andi *ССЫЛКА НА КАКУЮ-НИБУДЬ РАБОТУ АЛЕКСЕЕВА*. As argued in [@daniel18] there is some evidence that across the family, the second inanimate class (in cases where there is such a distinction) and similar-looking plural markers, ultimately originate in a distinct class containing mass nouns, such as liquids, pluralia and singularia tantum and abstract nouns that are "number neutral". If that is the case, the inanimate 2 has extended its terrain considerably in contemporary Andi. It contains some mass nouns and singularia tantum like *ilajma* 'parents' (literally a contraction of 'mother' (*ila*) and 'father' (*ima*), which agrees with inanimate 2 singular), but also many countable objects. In addition, mass nouns are not restricted to inanimate 2, for example *zar* 'ice' and *k'ːöj* 'smoke' belong to inanimate 1 at least in Rikvani. Thus, evidence from Andi does not necessarily refute the mass noun hypothesis, but it cannot support it either, as we have no knowledge of how Andi noun classes have evolved. Moreover because we have not yet succeeded in establishing semantic regularities for the synchronic level.

#### 3.3 Loan words
